---
title: "Time on the Horizon: ggHoriPlot methods"
date: "3/6/2023"
date-modefied: "`r Sys.Date() `"
format: html
execute: 
  echo: true
  eval: true
  warning: false
  freeze: auto
---

## Getting Started - Load packages and import data

Before getting start, make sure that ggHoriPlot has been included in the pacman::p_load(...) statement above.

```{r}
#| code-fold: true
pacman::p_load(ggHoriPlot, ggthemes, tidyverse) #no need to include lubridate here because lubridate is part of tidyverse library
```

Use the code chunk below to import the AVERP.csv file into R environment. The data file contains data on the Average Retail Prices of Selected Consumer Items. 

```{r}
#| code-fold: true
averp <- read_csv("data/AVERP.csv") %>%
  mutate(`Date` = dmy(`Date`)) #need to mutate because the original data is in chr format
```

## Plotting the horizon graph

The code chunk below will be used to plot the horizon graph.

```{r}
#| code-fold: true

averp %>% 
  filter(Date >= "2018-01-01") %>% #this line filters specific dates of interest
  ggplot() + #calls ggplot, necessary since the horizon plot is an extension of ggplot
  geom_horizon(aes(x = Date, y=Values), 
               origin = "midpoint", 
               horizonscale = 6)+ #indicating 6 months interval
  facet_grid(`Consumer Items`~.) + #use a ` because there is a space in consumer items
    theme_few() + #codes after this line is cosmetic
  scale_fill_hcl(palette = 'RdBu') +
  theme(panel.spacing.y=unit(0, "lines"), strip.text.y = element_text(
    size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
    scale_x_date(expand=c(0,0), date_breaks = "6 month", date_labels = "%b%y") +
  ggtitle('Average Retail Prices of Selected Consumer Items (Jan 2018 to Dec 2022)')

```

```{r}
#| code-fold: true

averp %>% 
  filter(Date >= "2018-01-01") %>% #this line filters specific dates of interest
  ggplot() + #calls ggplot, necessary since the horizon plot is an extension of ggplot
  geom_horizon(aes(x = Date, y=Values), 
               origin = "midpoint", 
               horizonscale = 6)+ #indicating 6 months interval
  facet_grid(`Consumer Items`~.) + #use a ` because there is a space in consumer items
    theme_few() + #codes after this line is cosmetic
  scale_fill_hcl(palette = 'RdBu') 
```

